---
title: "Rocket League Sideswipe Mystery Items"
author: "Toby Hayward"
date: "2022-10-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Summary

This is a document outlining the 91 mystery item crates that I opened. I want to use this data to estimate the probabilities of each item and provide visualisations to convey these findings.


```{r data}
data = read_csv('data/rlssmysteryboxes.csv')

```

A brief investigation to make sure that I wasn't too asleep when I collected this data.

```{r data quality check 1}
data %>% 
  ggplot(aes(x = quality, y = quality.colour)) +
  geom_tile()
```

Okay looks like I got the qualities right every time. 

```{r data quality check 2}
items = data %>% 
  group_by(item, quality) %>% 
  count() %>% 
  pull(item)

items

length(items) / length(unique(items))

```

This gives me reason to assume that I didn't miscategorise any item.\
I'm pretty satisfied with the quality at this point, I'm just going to tidy the data a bit.

```{r data tidy}
data.tidy = data %>%
  mutate(
    # Reorder the qualities. This link https://theglobalgaming.com/rocket-league/rarity-chart#:~:text=Common%3A%20Complete,are%20extremely%20rare.
    quality = factor(toupper(quality), levels = toupper(c('common', 'uncommon', 'rare', 'very rare', 'import', 'exotic', 'black market'))),
    item = toupper(item)
  )

data.tidy


```


Now for some questions. The questions I am interested in are:

* What is the expected return on value for a mystery box purchase?
* What are the estimated probabilities of each quality?
* What **item type** is the most frequent?
* How common are **coloured** items?
* What is the distribution of item types for each quality?

Let's start with the second question since we'll use it to aid the first: *What is the estimated probablilities of each quality?*

```{r probabilities}
# Get the colours of the qualities
quality.colours = data.tidy %>% 
  group_by(quality) %>% 
  slice_head() %>% 
  # Conveniently in the correct order. It seems like the function did this intentionally judging by the order of box.price.
  pull(quality.colour)


# Those colours are wack. Let's choose nice ones from http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf

quality.colours.decent = c(
  'grey90', # can't deny it
  'cyan2',
  'dodgerblue',
  'darkorchid1',
  'red',
  'gold',
  'deeppink2'
)

# This is better

data.tidy %>% 
  ggplot(aes(x = quality, fill = quality)) +
  geom_bar(col = 'black') +
  scale_fill_manual(values = quality.colours.decent) +
  theme_dark() +
  labs(title = 'Frequencies of Item Type',
       subtitle = paste('From a sample of', nrow(data), 'Mystery Items'),
       caption = 'Taken on 30/10/2022',
       x = 'Item Quality', y = 'Frequency') +
  theme(plot.title = element_text(size = 30, face = 2, hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = 'none',
        axis.line = element_line(colour = 'black'))
```

```{r probabilities 2}

# Now for the official proportions

frequencies = data.tidy$quality %>% table()
(probabilities.quality = frequencies / sum(frequencies))

data.tidy %>% 
  group_by(quality) %>% 
  summarise(p = n() / nrow(data)) %>% 
  ggplot(aes(x = quality, fill = quality, y = p * 100)) +
  geom_col(col = 'black') +
  scale_fill_manual(values = quality.colours.decent) +
  theme_dark() +
  labs(title = 'Probabilities of Item Type',
       subtitle = paste('From a sample of', nrow(data), 'Mystery Items'),
       caption = 'Taken on 30/10/2022',
       x = 'Item Quality', y = 'Probability %') +
  theme(plot.title = element_text(size = 30, face = 2, hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = 'none',
        axis.line = element_line(colour = 'black')) +
  scale_y_continuous(breaks = seq(0, 40, 10), labels = c(paste0(seq(0, 40, 10), '%')))
```



Judging by the *current* item shop. It appears that each quality has the following value;

* Common ~ No Value
* Uncommon Topper ~ 400 SP
* Rare Banner ~ 800 SP
* Very Rare Boost ~ 1200 SP 
* Import Boost ~ 1800 SP
* Exotic Wheel ~ 2400 SP
* Black Market Car Body ~ 3200 SP

And these videos [1](https://www.youtube.com/watch?v=aIi6ypdDf-E) [2](https://www.youtube.com/watch?v=6gic-dX-EwY). 

* Common ~ No Value
* Uncommon Banner ~ 400 SP
* Uncommon Decal ~ 400 SP
* Rare Decal ~ 800 SP
* Rare Title ~ 800 SP
* Very Rare Boost ~ 1200 SP 
* Very Rare Wheel ~ 1200 SP 
* Import Boost ~ 1800 SP
* Exotic Wheel ~ 2400 SP
* Exotic Car Body ~ 2400 SP
* Black Market Car Body ~ 3200 SP
* Black Market Goal Explosion ~ 3200 SP

Because we can't let items have no value, let's just assume that *COMMON* items have **200** SP in value. Hence it seems like the estimated values follow a trend shown below. 

```{r quality values, message = FALSE}
data.quality.value = tibble(quality = factor(levels(data.tidy$quality), levels = levels(data.tidy$quality)),
       value = c(200, 400, 800, 1200, 1800, 2400, 3200))
data.quality.value

data.quality.value %>% 
  ggplot(aes(x = quality, y = value)) +
  geom_point(col = quality.colours.decent) +
  geom_smooth(aes(x = 1:7), col = alpha('red', 0.2), alpha = 0.2, se = F) +
  geom_point(col = quality.colours.decent) +
  theme_bw() +
  labs(title = 'Quality Value',
       subtitle = 'As Estimated by the Item Shop on 3 Occasions',
       x = 'Quality', y = 'Value (SP)')

```



Therefore now we can use this data to answer the question: *What is the expected return on value for a mystery box purchase?*

```{r expected return}
(data.quality.value.prob = data.quality.value %>% 
  left_join(tibble(
    quality = names(probabilities.quality),
    prob = probabilities.quality %>% as.numeric()
  ), by = 'quality') %>% 
  mutate(exp_value = value * prob)) 

data.quality.value.prob %>% 
  pull(exp_value) %>% sum()

```


Therefore, you are always likely to gain value when opening a mystery box. This raises an important question though, *How likely?*... \
We can answer this in two ways: 

1. Approximate the distribution and calculate it's probability (Long and boring)
2. Just use the data and assume it's representative and sufficient (which it very may well be).


```{r how likely lol p1}
(data.tidy2 = data.tidy %>% 
  left_join(data.quality.value))

data.tidy2 %>% 
  ggplot(aes(x = value)) +
  geom_bar() +
  theme_bw()

data.tidy2 %>% 
  ggplot(aes(x = value)) +
  geom_density() +
  theme_bw()

data.tidy2 %>% 
  ggplot(aes(x = log(value))) +
  geom_bar() +
  theme_bw()

data.tidy2 %>% 
  ggplot(aes(x = log(value))) +
  geom_density() +
  theme_bw()


```


This looks like a *Quasi-Poisson* Distribution. We can approximate it's parameters using *Maximum Likelihood.* \
Because Quasi-Poisson does not have a *Probability-Density-Function*, instead we can transform the data 


```{r mle}
# Screw that! For now...
plot(0:12, dpois(0:12, 6))

```


The second way I can calculate it is just by estimating the probability of getting something below 500 SP. 

```{r crude way! use the data dummy}
(probability.profit = (data.tidy2$value[data.tidy2$value > 500] %>% length()) / length(data.tidy2$value))

```

My data claims that if you open a crate, you have a `r round(probability.profit, 3) * 100`% chance of getting an item that is worth more than the investment into the crate.










